---
title: "DigiCAT Counterfactual Analysis Report"
output: pdf_document
params:
  data_name: NA
  data: NA
  outcome_variable: NA
  treatment_variable: NA
  matching_variables: NA
  covariates: NA
  weighting_variable: NA
  non_response_weight: NA
  cluster_variable: NA
  stratification_variable: NA
  CF_approach: NA
  missingness: NA
  balancing_model: NA
  matching_method: NA
  matching_ratio: NA
  common_support_plot: NA
  observation_table: NA
  love_plot: NA
  balance_table: NA
  outcome_formula: NA
  outcome_variable_type : NA
  outcome_res: NA
  include_sensitivity : NA
  sensitivity_results : NA
editor_options: 
  markdown: 
    wrap: 72
---

## Sample used in counterfactual analysis

Data from `r params$data_name` was used for this analysis. This data set
contains `r nrow(params$data)` rows and `r ncol(params$data)` columns. A
total of `r nrow(params$data) - nrow(na.omit(params$data))` rows
containing missing data were detected.

## Input variables

Counterfactual analysis was used to measure the potential causal effect
of `r params$treatment_variable` on `r params$outcome_variable`.
Matching variables included
`r sub(",([^,]*)$"," and\\1", paste0(params$matching_variables, collapse = ", "))`.

```{r input, echo=FALSE, include=TRUE, comment=NA}

additional_input_variables <- NULL

## If covariates exist, name them
if (!is.null(params$covariates)){
  additional_input_variables <- c(additional_input_variables, paste0("Additional covariates included ", sub(",([^,]*)$"," and\\1", paste0(params$covariates, collapse = ", "))))
}

## If survey weights exist, name
if (!is.null(params$weighting_variable) & params$non_response_weight == FALSE){
  additional_input_variables <- c(additional_input_variables, paste0("Survey weights were indicated in ", params$weighting_variable, "."))
}

if (!is.null(params$weighting_variable) & params$non_response_weight == TRUE){
  additional_input_variables <- c(additional_input_variables, paste0("Survey weights were indicated in ", params$weighting_variable, ". These weights compensated for non-response."))
}

## If cluster exist, name
if (!is.null(params$cluster_variable)){
  additional_input_variables <- c(additional_input_variables, paste0(params$cluster_variable, " indicated clusters within the sample."))
}

## If stratification exist, name
if (!is.null(params$stratification_variable)){
 additional_input_variables <- c(additional_input_variables, paste0(params$stratification_variable, " was used to stratify the sample."))
}

additional_input_variables <- paste(additional_input_variables, collapse = " ")
```
`r additional_input_variables`

## Counterfactual Approach
```{r cf approach, include = FALSE}

CF_approach <- NULL

## Counterfactual approach
## PSM
if(params$CF_approach == "psm"){
  CF_approach <- paste0("Propensity score matching was used to carry out counterfactual analysis.")
}

## IPTW
if(params$CF_approach == "iptw"){
  CF_approach <- paste0(CF_approach, paste0("Inverse probability of treatment weighting (IPTW) was used to carry out counterfactual analysis."))
}

## NBP
if(params$CF_approach == "nbp"){
   CF_approach <- paste0(CF_approach, paste0("Non-bipartite propensity matching (NBP) was used to carry out counterfactual analysis."))
}

## CBPS
if(params$CF_approach == "cbps"){
   CF_approach <- paste0(CF_approach, paste0("Covariate balancing propensity score (CBPS) was used to carry out counterfactual analysis."))
}


## Data missingness
## If no missing data
if (nrow(params$data) - nrow(na.omit(params$data)) == 0){
  CF_approach <- paste0(CF_approach, paste0(" No missing data was detected."))
}

## MI
if ((nrow(params$data) - nrow(na.omit(params$data)) > 0) & params$missingness == "mi"){
  CF_approach <- paste0(CF_approach, paste0(" Missing data was handled using multiple imputation (m = 5)."))
}

## CC
if ((nrow(params$data) - nrow(na.omit(params$data)) > 0) & params$missingness == "complete"){
  CF_approach <- paste0(CF_approach, paste0(" Missing data was handled using complete case analysis (CCA)."))
}

## weighting
if ((nrow(params$data) - nrow(na.omit(params$data)) > 0) & params$missingness == "weighting"){
  CF_approach <- paste0(CF_approach, paste0(" Missing data was handled using non-response weighting."))
}

## Balancing model
if (!is.null(params$balancing_model)){  ## Only run if balancing model given
  if (params$balancing_model == "glm"){
    CF_approach <- paste0(CF_approach, paste0(" A logistic regression model was used to calculate propensity scores in the sample."))
  }
  if (params$balancing_model == "gbm"){
    CF_approach <- paste0(CF_approach, paste0(" Generalized boosted model (GBM) was used to calculate propensity scores in the sample."))
  }
  if (params$balancing_model == "rf"){
    CF_approach <- paste0(CF_approach, paste0(" Random forest was used to calculate propensity scores in the sample."))
  }
  if (params$balancing_model == "orl"){
    CF_approach <- paste0(CF_approach, paste0(" An ordinal logistic regression was used to calculate propensity scores in the sample."))
  }
  CF_approach <- paste(CF_approach, collapse = " ")
}

```
`r CF_approach`

## Balancing
```{r balancing, echo=FALSE}
balancing <- NULL

## Balancing parameters

if(params$CF_approach == "psm"){
  balancing <- paste0("Matching was carried out using a ratio of 1:", params$matching_ratio, " and a matching method of '", params$matching_method,"'. ")
}

if(params$CF_approach == "iptw"){
  balancing <- paste0("In order to balance the sample across treatment groups, the propensity scores estimates are transformed to provide weights to be used in the outcome model.")
}

if(params$CF_approach == "nbp"){
  balancing <- paste0("NBP uses 1:1 optimal matching, minimising the overall propensity score differences between matched cases. The algorithm in DigiCAT also disallows matches of pairs that are too dis-similar to help ensure that there is good balance between the groups after matching.")
}

if(params$CF_approach == "cbps"){
  balancing <- paste0("CBPS was applied to optimize balance between treatment groups while simultaneously estimating propensity scores. This method improves covariate balance by incorporating balance constraints directly into the propensity score estimation process.")
}

balancing <- paste0(balancing, paste0("Balancing output can be seen below."))

balancing <- paste(balancing, collapse = " ")
```

`r balancing`

\newpage

```{r common support plot, echo=FALSE, include=TRUE, results='asis'}
## Start plot number counter
plot_number <- 1

## Common support graph
if(!is.null(params$common_support_plot)){
  plot_number <- plot_number + 1
  ## Print the plot
  print(params$common_support_plot)
  ## Output plot caption
  description <- paste0("Plot ", plot_number, ": Common Support Graph")
  cat("\n", description, "\n")
  ## Insert a new page
  cat("\n\\newpage\n")
}

```



```{r observation table , echo=FALSE, include=TRUE, comment=NA}

## Observation table
knitr::kable(params$observation_table, caption = "Observation table", format="latex", 
      align = c("|p{6cm}","p{6cm}","p{6cm}|"), position = "h!")
```

\newpage

```{r love plot, echo=FALSE, include=TRUE, results='asis'}

## Print the plot
plot_number <- plot_number + 1
## Print the plot
print(params$love_plot)
## Output plot caption
description <- paste0("Plot ", plot_number, ": Love Plot")
cat("\n", description, "\n")
## Insert a new page
cat("\n\\newpage\n")
```

\newpage

```{r balance table, echo=FALSE, include=TRUE, comment=NA}

## Balance table
if (params$missingness == "mi"){ ## More columns with MI

knitr::kable(params$balance_table, caption = "Balance table", format="latex", 
      align = c("|p{1.5cm}","p{1.5cm}", "p{1.5cm}", "p{1.5cm}", "p{1.5cm}", "p{1.5cm}","p{1.5cm}|"), position = "h!")
} else{
  knitr::kable(params$balance_table, caption = "Balance table", format="latex", 
      align = c("|p{4cm}","p{4cm}", "p{4cm}","p{4cm}|"), position = "h!")

}

```

\newpage

## Outcome Model
```{r outcome, echo = FALSE}
outcome_model <- NULL

if(params$outcome_formula == "unadjusted"){
  
  if(params$outcome_variable_type == "continuous"){
      outcome_model <- paste0("Linear regression without matching variables as covariates was used as the outcome model in this analysis.")
  }   
  if(params$outcome_variable_type == "binary"){
          outcome_model <- paste0("Logisitic regression without matching variables as covariates was used as the outcome model in this analysis.")
  }  
  if(params$outcome_variable_type == "categorical"){
              outcome_model <- paste0("Multinomial logistic regression without matching variables as covariates was used as the outcome model in this analysis.")
  }
}

if(params$outcome_formula == "with_matching_variables"){
  
  if(params$outcome_variable_type == "continuous"){
      outcome_model <- paste0("Linear regression adjusted for matching variables as covariates was used as the outcome model in this analysis.")
  }   
  if(params$outcome_variable_type == "binary"){
          outcome_model <- paste0("Logisitic regression adjusted for matching variables as covariates was used as the outcome model in this analysis.")
  }  
  if(params$outcome_variable_type == "categorical"){
              outcome_model <- paste0("Multinomial logistic regression adjusted for matching variables as covariates was used as the outcome model in this analysis.")
  }
}

if(params$outcome_formula == "marginal_effects"){
  
    if(params$outcome_variable_type == "continuous"){
      outcome_model <- paste0("Linear regression adjusted for matching variables as covariates and their interactions with the treatment variable was used as the outcome model of this analysis.")
  }   
  if(params$outcome_variable_type == "binary"){
          outcome_model <- paste0("Logisitic regression adjusted for matching variables as covariates and their interactions with the treatment variable was used as the outcome model of this analysis.")
  }  
  if(params$outcome_variable_type == "categorical"){
              outcome_model <- paste0("Multinomial logistic regression adjusted for matching variables as covariates and their interactions with the treatment variable was used as the outcome model of this analysis.")
  }
}

outcome_model <- paste0(outcome_model," The estimated effect of ", params$treatment_variable, " on ", params$outcome_variable, " was ", round(params$outcome_res[1,"Coefficient Estimate"], 4), "(p-value = ", round(params$outcome_res[1,"P-value"], 4), ", standard error = ", round(params$outcome_res[1,"Standard Error"], 4), ", 95% confidence interval:", round(params$outcome_res[1,"Lower CI (2.5%)"], 4), " to ", round(params$outcome_res[1,"Upper CI (97.5%)"], 4),  ")" )

```

`r outcome_model`

\newpage

```{r sensitivity analysis, echo=FALSE, include=TRUE, results='asis'}
sensitivity_analysis <- NULL

if (params$include_sensitivity == TRUE){
  
  cat("## Sensitivity Analysis")
  
  cat("\nTo assess the robustness of our findings to unmeasured confounding, we conducted sensitivity analysis. Rosenbaum’s bounds assess how hidden bias may affect study results using a sensitivity parameter, Gamma. If p-values stay significant at higher Gamma values, results are robust; otherwise, they may be sensitive to hidden bias. DigiCAT computes Gamma from X to Y.\n\n")
  
  
 print(knitr::kable(params$sensitivity_results$RB$bounds, caption = params$sensitivity_results$RB$msg, format="latex") %>%
    kable_styling(latex_options = "hold_position"))
  cat("\n")
  cat("\n", params$sensitivity_results$RB$note, "\n")
  
  if(!is.null(params$sensitivity_results$EV)){

  cat("\nE-values estimate the strength an unmeasured confounder must have to negate an observed association. High E-values suggest robust findings\n")
     
  print(knitr::kable(temp$sensitivity_results$EV, caption = "E-Values", format="latex") %>%
    kable_styling(latex_options = "hold_position"))
    
  }
  
  ## Insert a new page
  cat("\n\\newpage\n")
}


```


## R package version information:

```{r session info, echo=FALSE, include=TRUE, }
package_versions <- as.data.frame(installed.packages()[,c(1,3:4)])
rownames(package_versions) <- NULL
package_versions <- package_versions[is.na(package_versions$Priority),1:2,drop=FALSE]

package_versions


```
